---
title: "Covid91 vaccine study Final2024"
author: "Racheal Adeneye"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="M")
gal = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```


# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

## Barplot (Men)
```{r}
dd2 <- guy %>%
  group_by(infected, treatment) %>%
  summarize(count = n(), .groups = "drop") %>%
  group_by(infected) %>%
  mutate(prcnt = count / sum(count))
```

# the group_by followed by summarize(count=n())
```{r}
basicC <- ggplot(dd2, aes(x = infected, y = count, fill = treatment))
print(basicC + geom_bar(stat = "identity", position = "dodge"))
```

## Percentage Plot (Men)
```{r}
basicCC <- ggplot(dd2, aes(x=infected, y=prcnt*100, fill=treatment))
print(basicCC + geom_bar(stat = "identity", position = "dodge"))
```

These plots display the population of male participants in the clinical trial and how their likelihood of contracting COVID-19 varied depending on whether they received the vaccine or a placebo.

The first plot shows that about 74% of participants who received the placebo contracted COVID-19, while only about 26% of vaccinated participants did. This indicates a significantly higher infection rate in the placebo group, suggesting the vaccine is effective in reducing the risk of COVID-19 infection. Based on the data, the odds of contracting COVID-19 after vaccination are approximately 1 in 3 compared to those who received the placebo.

In the first plot, the x-axis shows whether participants were infected or not, and the y-axis shows the total count of individuals in each category.

In the second plot, the x-axis still represents COVID-19 infection status, but the y-axis now represents the percentage of individuals who contracted the virus, categorized by whether they received the vaccine or the placebo. The two left bars represent all the individuals who got infected.

The data used to create these plots come from different objects in the environment pane. The object dd2 contains four variables: infected, treatment, count, and prcent (short for "percent"). This dataset contributes to the second, percentage-based plot. The basicC object includes nine variables and is used to create the first plot. Another object, basicCC, also contains nine variables and contributes to generating the second plot as well.

### Numerical results (Men)

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```
These numerical results highlight the specific percentages of male participants who contracted COVID-19 based on whether they received the vaccine or the placebo. About 26% of those who were vaccinated became infected, compared to approximately 74% of those who received the placebo. This demonstrates that the vaccine is highly effective in preventing COVID-19 amongst men.

### Inferential results (Men)
```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```
The key point to focus on is the output from Fisher’s Exact Test. The reported p-value is extremely small, 0.0000000000000002, indicating that the observed results are highly unlikely to have occurred by chance.

The null hypothesis states that COVID-related deaths are independent of sex. However, the data strongly suggests otherwise, supporting the alternative hypothesis: that there is a significant association between sex and COVID-19 death rates.

Additionally, the odds ratio is about 2.82, meaning that individuals in the placebo group were about 2.82 times more likely to contract COVID-19 compared to those who received the vaccine.

## Effectiveness Confidence Interval (Men)

```{r}
effect <- guysfish$p.value/(guysfish$p.value +1)
effectivenessguy = guysfish$estimate/(guysfish$estimate +1)*100
print(effectivenessguy)
```
```{r}
low=(guysfish$conf.int[1]/(guysfish$conf.int[1] +1))*100
hi=(guysfish$conf.int[2]/(guysfish$conf.int[2] +1))*100
options(digits=2)
```

We can see above the effectiveness of the COVID-19 vaccine in male participants. An effectiveness rate above 50% indicates that the vaccine offers protection. Based on the confidence interval displayed, we can conclude that the vaccine is effective in reducing the risk of infection among men.

The odds ratio provides an estimate of how effective the vaccine is. For males, the odds ratio is approximately 74, with a confidence interval ranging from 54.73 to 72.32, further supporting the vaccine’s strong protective effect.



## Females

## Barplot (Women)
```{r}
dd2 <- gal %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
```

# the group_by followed by summarize(count=n())
```{r}
basicC <- ggplot(dd2,aes(x = infected,y = count,fill = treatment))
basicC + geom_bar(stat = "identity", position = "dodge")
```

## Percentage Plot (Women)
```{r}
basicCC <- ggplot(dd2,aes(x = infected,y = prcnt*100,fill = treatment)) 
basicCC + geom_bar(stat = "identity", position = "dodge")
```
This plot compares COVID-19 infection outcomes among female participants who received either the vaccine or the placebo.

Similar to the previous plot, about 65% of the females who received the placebo contracted COVID-19, which is a notably high rate. This means that roughly 65% of the infected individuals were in the placebo group. The odds of contracting COVID-19 while on the placebo were 2.8 times higher, compared to about 1.02 times for those who received the vaccine. In other words, women who were vaccinated were less likely to become infected than those who took the placebo.

There are two plots shown. In the first plot, the x-axis indicates whether participants were infected with COVID-19 or not, while the y-axis shows the count of individuals in each category.

In the second plot, the two leftmost bars represent all participants who contracted COVID-19. The x-axis again shows infection status, but the y-axis now reflects the percentage of people infected, grouped by whether they took the vaccine or the placebo.

The code used to generate these plots starts with the block labeled dd2, visible in the environment pane. This dataset contains four variables: infected, treatment, count, and percent. Another code group, basicC, also in the environment pane, contains nine variables and is used to create the first plot. Finally, basicCC, found in the files pane with nine variables, helps produce the second, percentage-based plot.

## Numerical Results (Women)
```{r}
table1 <- xtabs(~infected + treatment, data=gal)
rowPerc(table1)
colPerc(table1)
```
These numerical results reveal the specific percentages of women who contracted COVID-19 after receiving either the placebo or the vaccine. About 35% of the women who were vaccinated became infected, consistent with the earlier plots. In contrast, 65% of the women who received the placebo contracted COVID-19. This clearly shows that the vaccine is effective in protecting women from the virus. Women who were vaccinated were significantly less likely to contract COVID-19 compared to those who took the placebo.


## Inferential Results (Women)
```{r}
chisq.test(table1)
chisqtestGC(table1)
galsfish <- fisher.test(table1)
galsfish
```
The key result to focus on is the output of Fisher’s Exact Test. The p-value shown above is extremely small, 0.00000001, indicating that these results are highly unlikely to have occurred by chance. Therefore, we can reject the null hypothesis in favor of the alternative.

Fisher’s Exact Test also shows an odds ratio of 1.9. Women who received the vaccine were about 1.9 times as likely to contract COVID-19 compared to those who received the placebo.

## Effectiveness Confidence Interval (Women)
```{r}
effect <- galsfish$p.value/(galsfish$p.value +1)
effectivenessgal = galsfish$estimate/(galsfish$estimate +1)*100
print(effectivenessgal)
```

```{r}
low=(galsfish$conf.int[1]/(galsfish$conf.int[1] +1))*100
hi=(galsfish$conf.int[2]/(galsfish$conf.int[2] +1))*100
options(digits=2)
```
The chart above shows the effectiveness of the COVID-19 vaccine in females. An effectiveness rate above 50% indicates that the vaccine is effective. Based on the confidence interval displayed, we can conclude that the vaccine is highly effective in protecting women from contracting COVID-19.

Generally, a vaccine is considered effective if its effectiveness is 60% or higher. This vaccine performs exceptionally well in preventing COVID-19 infections among women.

The odds ratio provides an estimate of the vaccine’s effectiveness. For females, the odds ratio is approximately 65, with a confidence interval ranging from 68.75 to 78.26, further confirming the vaccine’s strong protective effect.




## LGBTQ

### Barplot (LGBTQ)

```{r}
dd2 <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
```

# the group_by followed by summarize(count=n())
```{r}
basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
```

## Percentage Plot (LGBTQ)
```{r}
basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This plot illustrates the COVID-19 infection outcomes among individuals who identify as gay, comparing those who received the vaccine with those who received a placebo.

According to the barplot, approximately 20% of vaccinated individuals contracted COVID-19, while about 80% of those who received the placebo became infected. This suggests that individuals in this group who were vaccinated were about 16 times less likely to get COVID-19 than those who were given the placebo. The odds of contracting the virus while vaccinated are roughly 1 in 4.5.

There are two plots displayed. In the first plot, the x-axis indicates whether participants were infected or not, and the y-axis shows the number of individuals in each category. In the second plot above, the x-axis still reflects infection status, but the y-axis now represents the percentage of individuals who got COVID-19 after receiving either the placebo or the vaccine. The two bars on the left highlight the infected portion of each group.

The plots are generated using code that begins with the object named dd2, which is visible in the environment pane and includes four variables: infected, treatment, count, and percent. The second code block, labeled basicC, also appears in the environment and contains nine variables that contribute to generating the first plot. The final code block, basicCC, contains another set of nine variables and is used to create the second plot—the one based on percentages.

### Numerical Results (LGBTQ)

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```
These numerical results highlight the percentage of gay individuals who contracted COVID-19 after receiving either the vaccine or the placebo. Among those who were vaccinated, 20% became infected, while 80% of those who received the placebo contracted the virus. These findings indicate that the vaccine was highly effective in protecting individuals in the gay community against COVID-19.

The odds of contracting COVID-19 while vaccinated were approximately 1 in 4.5, meaning vaccinated individuals in this group were significantly less likely to become infected compared to those who received the placebo.

## Inferential Results (LGBTQ)
```{r}
chisq.test(table1)
chisqtestGC(table1)
gaysfish <- fisher.test(table1)
gaysfish
```
The p-value is 0.0000007, an extremely small number, indicating that these results are highly unlikely to have occurred by chance.

Fisher’s Exact Test shows an odds ratio of 4.3, meaning that gay individuals who received the vaccine were approximately 77% less likely, or only 0.23 times as likely, to contract COVID-19 compared to those who received the placebo.

Given this evidence, we reject the null hypothesis in favor of the alternative. The vaccine appears to be highly effective in reducing the likelihood of COVID-19 infection among individuals who identify as gay.

## Effectiveness Confidence Interval (LGBTQ)
```{r}
effect <- gaysfish$p.value/(gaysfish$p.value +1)
effectivenessgays = gaysfish$estimate/(gaysfish$estimate +1)*100
print(effectivenessgays)
```
```{r}
low=(gaysfish$conf.int[1]/(gaysfish$conf.int[1] +1))*100
hi=(gaysfish$conf.int[2]/(gaysfish$conf.int[2] +1))*100
options(digits=2)
```

The graph above shows the effectiveness of the COVID-19 vaccine among individuals who identify as gay. An effectiveness rate above 50% indicates that the vaccine is considered effective. Based on the confidence interval displayed, we can conclude that the vaccine offers strong protection for this group against contracting COVID-19.

The odds ratio, which provides a measure of the vaccine’s effectiveness, is shown above as 81, indicating a substantial level of protection. This suggests that vaccinated individuals are significantly better protected compared to those who did not receive the vaccine. The confidence interval for vaccine effectiveness in this group ranges from 71.4% to 86.7%.



## Druggies


# Overall Results and Conclusions